### 2013-07-18 Tex/chapters/results.tex
kleine Umformulierung
Reihenfolge verändert

\subsection{Finite Size Effects}
\label{ssec:finitesize}
    The aim of this thesis is to find the critical temperature \(T_c\)
    of the disordered Ising model in dependence of the disorder parameter
    \(\sigma\). At \(T_c\) the mean absolute magnetization \(\avg{|m|}\) of
    the system will jump and the susceptibility
    \(\chi = \frac{N}{T}\brac{\avg{m^2}-\avg{m}^2}\)
    will diverge. In fig. \ref{fig:smeared_out}\subref{sfig:smeared_out:meanM_L}
    it is easy to see that the jump of \(\avg{|m|}\)
    occurs at lower temperatures \(T\) for higher
    disorder parameters \(\sigma\) (here with a RNG).
        \subfigure[Dependence of the phase transition on $\sigma$][]
            \label{sfig:smeared_out:meanM_L}
            \includegraphics[width=0.45\textwidth]{plots/Mean_M_L_128}
        \subfigure[Example for finite size effects][]
            \label{sfig:smeared_out:meanM}
            \includegraphics[width=0.45\textwidth]{plots/meanM}
        \caption[Phase Transition and Finite Size Effects]
            \subref{sfig:smeared_out:meanM_L}: Effects of the disorder
            parameter $\sigma$ on the phase transition
            for an underlying RNG with $L=128$.\\
            \subref{sfig:smeared_out:meanM}: Effects of different system
            sizes at \(\sigma = 0\). Dotted lines are guides to the eye.
        \label{fig:smeared_out}
    But in the shown plots there occurs no real jump, but a continuous
    decline. The jump is only present in infinite systems, hence no
    computer simulation will show the exact behavior in the
    thermodynamic limit. It will always show some \emph{finte size effects}.
    These finite size effects cause a "smearing out" of the phase
    transition. This is stronger for smaller system sizes, as is visible
    in fig. \ref{fig:smeared_out}\subref{sfig:smeared_out:meanM}\footnote{See the appendix \ref{appendix:finiteSizeEffects} for a similar figure for the specific heat.}.
    Clearly the \(L=16\) curve is much less steep than the \(L=128\) curve.\\
    Despite of this one can obtain \(T_c\) by finite size scaling
    methods \cite[S. ??]{NewmanBarkema1999} which also yield the critical
    exponents.
    Studies on random lattices with\cite{Lima2000} and without\cite{Janke1994}
    varying coupling constants \(J\) suggest, that the
    critical exponents should not be influenced by the disorder parameter
    \(\sigma\). So they will be used for consistency cross checking and
    comparison with the known exact values \cite[S. 59]{Pelissetto2002}.\\
    % Ausfuehrlichere Herleitung der Skalentheorie, Skalenanalyse
    For a more in depth look at finite size scaling, \cite{Norrenbrock2011}
    offers an explanation.
    It is known, that for large \(L\) near the critical temperature the
    equations \eqref{eq:fsscaling:m}, \eqref{eq:fsscaling:chi} and
    \eqref{eq:fsscaling:g} apply.
    \begin{align}
        \label{eq:fsscaling:m}
        \avg{m_L} &= L^\frac{\beta}{\nu} \tilde{M}\brac{L^\frac{1}{\nu}\brac{T-T_c}}\\
        \label{eq:fsscaling:chi}
        \chi_L    &= L^\frac{\gamma}{\nu} \tilde{C}\brac{L^\frac{1}{\nu}\brac{T-T_c}}\\
        \label{eq:fsscaling:g}
        g          &= \frac{3}{2}\brac{1-\frac{\avg{m^4}}{3\avg{m^2}^2}} \propto \tilde{G}\brac{L^\frac{1}{\nu}\brac{T-T_c}}
    \end{align}
    Where \(g\) from eq. \eqref{eq:fsscaling:g} is a normalized
    Binder cumulant \cite{Binder1981} and \(\tilde{M}, \tilde{C}\) and \(\tilde{G}\)
    are unknown scaling functions. To find the exponents
    \(\beta, \gamma, \nu\) and the critical temperature \(T_c\), one
    varies them until the measured values of the observables collapse on
    one curve. Like the obersables from fig. \ref{fig:gettingCrit}\subref{sfig:gettingCrit:binder_fit_s_0}
    collapse in fig. \ref{fig:gettingCrit}\subref{sfig:gettingCrit:collapse_s_0}%\footnote{Examples for a collapse to determine the other critical exponents are in appendix \ref{appendix:finiteSizeScaling}}.
    Note that \(L=16\) is not used for the collapse, because it is a
    rather small value. Eq. \eqref{eq:fsscaling:m}-\eqref{eq:fsscaling:g}
    are approximations for big \(L\). For small \(L\) one needs some
    \emph{corrections to scaling} terms, which are not considered here.\\
    To accomplish the collapse in an semi-automatic and reproduceable
    way with an error estimate, the program
    \texttt{autoscale.py} \cite{autoscale2009} is used.
        \subfigure[Example of a Binder cumulant to determine the critical temperature][]
            \label{sfig:gettingCrit:binder_fit_s_0}
        \subfigure[Example of a datacollapse to determine critical exponents][]
            \label{sfig:gettingCrit:collapse_s_0}
            \includegraphics[width=0.47\textwidth]{plots/collapse_s_0}
            (\(\sigma=0\))\\
            \subref{sfig:gettingCrit:binder_fit_s_0} interpolated
                with cubic splines (the errorbars are too small to see)\\
            \subref{sfig:gettingCrit:collapse_s_0} collapsed by finite
                size scaling
        \label{fig:gettingCrit}
    \(g\) of different system sizes \(L\) because they are intersecting
    at \(T_c\) \cite{Binder1981}.
    As an example take fig. \ref{fig:gettingCrit}\subref{sfig:gettingCrit:binder_fit_s_0}.
    deviations.

\subsection{Critical Temperature}
\label{ssec:binderIntersections}
    The evaluation of the Binder cumulant's intersections, yields the
    critical temperatures \(T_c\).
    In this section these \(T_c\) on either the RNG or the GG are
    compared. In the following figures, the RNG will always be on the
    left side and the GG on the right side.
    In fig. \ref{fig:Tc} \(T_c\) is plotted over \(\sigma\).
    \begin{figure}[htbp]
        \centering
        \subfigure[][]
        {
            \label{sfig:Tc:RNG}
            \includegraphics[width=0.45\textwidth]{plots/RNG_Tc}
        }
        \subfigure[][]
        {
            \label{sfig:Tc:GG}
            \includegraphics[width=0.45\textwidth]{plots/GG_Tc}
        }
        \caption[Critical Temperature over different disorder parameters]
        {
            Critical temperatures \(T_c\) over different
            disorder parameters \(\sigma\) for
            \subref{sfig:Tc:RNG} the RNG and
            \subref{sfig:Tc:GG} the GG.\\
        }
        \label{fig:Tc}
    \end{figure}\\
    One sees that the RNG on the left has generally a lower critical
    temperature than the GG. Also \(T_c\) of the RNG decreases
    monotonically while \(T_c\) of the GG rises at first. Naively one would
    expect, that while changing the displacement of the nodes monotonically,
    the properties of the system will also change monotonically. The
    maximum at \(\sigma \approx 0.15\) on the GG will be discussed later.
    For big \(\sigma\) the displaced nodes approach the limit of randomly
    distributed nodes, hence \(T_c\) is independent of \(\sigma\) for
    \(\sigma \gg 1\).\\
    Another strange property is the jump from \(\sigma = 0\) to \(\sigma > 0\)
    of \(T_c\) on the GG. To understand that, one has to
    consider the influence of the graph properties on the critical
    temperature.\\
    One basic property of a graph is its degree \(K\) -- sometimes
    called \emph{coordination number} -- is defined as the mean count of
    neighbors per node
    \begin{equation}
        K = \frac{\sum_{\avg{i,j}} 1}{\sum_{i} 1}
        \label{eq:degree}
    \end{equation}
    It is known, that the degree has an impact on the critical Temperature.
    For example the Honeycomb lattice is of degree \(K=3\)
    and there exists an analytic solution \cite{Wannier1945}.
    Its critical temperature is determined by eq. \eqref{eq:exactHCTc}.
    \begin{equation}
        \cosh\brac{\frac{J}{T_c}}=2 \overset{J=1}{\Longrightarrow} T_c \approx 1.52
        \label{eq:exactHCTc}
    \end{equation}
    This is lower than for the square lattice with degree 4.\\
    If one plots the degree of the graphs at different \(\sigma\) like
    in fig. \ref{fig:Tc_deg}\subref{sfig:deg:RNG}\subref{sfig:deg:GG},
    one recognizes, that \(T_c\) and \(K\) are evidently correlated.
    It seems reasonable to normalize \(T_c\) by the degree of the underlying
    graph. This is done in fig \ref{fig:Tc_deg}\subref{sfig:Tc_norm_deg:RNG}\subref{sfig:Tc_norm_deg:GG}.
    Indeed the normalization eliminates the jump and reduces the
    slope of the rising on the GG. Hence it reduces differences
    between the RNG and the GG.
    \begin{figure}[htbp]
        \centering
        \subfigure[][]
        {
            \label{sfig:deg:RNG}
            \includegraphics[width=0.45\textwidth]{plots/RNG_deg}
        }
        \subfigure[][]
        {
            \label{sfig:deg:GG}
            \includegraphics[width=0.45\textwidth]{plots/GG_deg}
        }

\subsection{Critical Exponents}

@@

    expectations are met. Hence 5 values of \(\sigma\) are analyzed,
    which are chosen to represent every interesting region from fig. \ref{fig:Tc}\subref{sfig:Tc:RNG}\subref{sfig:Tc:GG}:
    the analytically known value for \(\sigma = 0\), the in \cite{Janke1994}
    examined limit for \(\sigma \gtrsim 1\), the plateau at small
    \(\sigma\), the steep decline and the shallow decline at \(\sigma = 0.5\).

@@

    are matching the expectations. Especially \(\nu\) and \(gamma\)

@@

    The results are displayed in fig. \ref{fig:susCrossCheck}. The by this
    method determined ratios confirm the by collapse obtained values.

@@

    disordered Ising model is for every \(\sigma\)in the same universality

@@

>>>>

\subsection{Finite Size Effects}
\label{ssec:finitesize}
    The aim of this thesis is to find the critical temperature \(T_c\)
    of the disordered Ising model in dependence of the disorder parameter
    \(\sigma\). At \(T_c\) the mean absolute magnetization \(\avg{|m|}\) of
    the system will jump and the susceptibility
    \(\chi = \frac{N}{T}\brac{\avg{m^2}-\avg{m}^2}\)
    will diverge. In fig. \ref{fig:smeared_out}\subref{sfig:smeared_out:meanM_L}
    it is easy to see that the jump of \(\avg{|m|}\)
    occurs at lower temperatures \(T\) for higher
    disorder parameters \(\sigma\) (here with a RNG).
    \begin{figure}[htbp]
        \centering
        \subfigure[Dependence of the phase transition on $\sigma$][]
        {
            \label{sfig:smeared_out:meanM_L}
            \includegraphics[width=0.45\textwidth]{plots/Mean_M_L_128}
        }
        \subfigure[Example for finite size effects][]
        {
            \label{sfig:smeared_out:meanM}
            \includegraphics[width=0.45\textwidth]{plots/meanM}
        }
        \caption[Phase Transition and Finite Size Effects]
        {
            \subref{sfig:smeared_out:meanM_L}: Effects of the disorder
            parameter $\sigma$ on the phase transition
            for an underlying RNG with $L=128$.\\
            \subref{sfig:smeared_out:meanM}: Effects of different system
            sizes at \(\sigma = 0\). Dotted lines are guides to the eye.
        }
        \label{fig:smeared_out}
    \end{figure}\\
    But in the shown plots there occurs no real jump, but a continuous
    decline. The jump is only present in infinite systems, hence no
    computer simulation will show the exact behavior in the
    thermodynamic limit. It will always show some \emph{finte size effects}.
    These finite size effects cause a "smearing out" of the phase
    transition. This is stronger for smaller system sizes, as is visible
    in fig. \ref{fig:smeared_out}\subref{sfig:smeared_out:meanM}\footnote{See the appendix \ref{appendix:finiteSizeEffects} for a similar figure for the specific heat.}.
    Clearly the \(L=16\) curve is much less steep than the \(L=128\) curve.\\
    Despite of this one can obtain \(T_c\) by finite size scaling
    methods \cite[S. ??]{NewmanBarkema1999} which also yield the critical
    exponents.

\subsection{Critical Exponents}
    Studies on random lattices with\cite{Lima2000} and without\cite{Janke1994}
    varying coupling constants \(J\) suggest, that the
    critical exponents should not be influenced by the disorder parameter
    \(\sigma\). So they will be used for consistency cross checking and
    comparison with the known exact values \cite[S. 59]{Pelissetto2002}.\\
    % Ausfuehrlichere Herleitung der Skalentheorie, Skalenanalyse
    For a more in depth look at finite size scaling, \cite{Norrenbrock2011}
    offers an explanation.
    It is known, that for large \(L\) near the critical temperature the
    equations \eqref{eq:fsscaling:m}, \eqref{eq:fsscaling:chi} and
    \eqref{eq:fsscaling:g} apply.
    \begin{align}
        \label{eq:fsscaling:m}
        \avg{m_L} &= L^\frac{\beta}{\nu} \tilde{M}\brac{L^\frac{1}{\nu}\brac{T-T_c}}\\
        \label{eq:fsscaling:chi}
        \chi_L    &= L^\frac{\gamma}{\nu} \tilde{C}\brac{L^\frac{1}{\nu}\brac{T-T_c}}\\
        \label{eq:fsscaling:g}
        g          &= \frac{3}{2}\brac{1-\frac{\avg{m^4}}{3\avg{m^2}^2}} \propto \tilde{G}\brac{L^\frac{1}{\nu}\brac{T-T_c}}
    \end{align}
    Where \(g\) from eq. \eqref{eq:fsscaling:g} is a normalized
    Binder cumulant \cite{Binder1981} and \(\tilde{M}, \tilde{C}\) and \(\tilde{G}\)
    are unknown scaling functions. To find the exponents
    \(\beta, \gamma, \nu\) and the critical temperature \(T_c\), one
    varies them until the measured values of the observables collapse on
    one curve. Like the obersables from fig. \ref{fig:gettingCrit}\subref{sfig:gettingCrit:binder_s_0}
    collapse in fig. \ref{fig:gettingCrit}\subref{sfig:gettingCrit:collapse_s_0}.
    %\footnote{Examples for a collapse to determine the other critical exponents are in appendix \ref{appendix:finiteSizeScaling}}.
    Note that \(L=16\) is not used for the collapse, because it is a
    rather small value. Eq. \eqref{eq:fsscaling:m}-\eqref{eq:fsscaling:g}
    are approximations for big \(L\). For small \(L\) one needs some
    \emph{corrections to scaling} terms, which are not considered here.\\
    To accomplish the collapse in an semi-automatic and reproduceable
    way with an error estimate, the program
    \texttt{autoscale.py} \cite{autoscale2009} is used.
    \begin{figure}[htbp]
        \centering
        \subfigure[Binder Cumulant $g$][]
        {
            \label{sfig:gettingCrit:binder_s_0}
            \includegraphics[width=0.47\textwidth]{plots/binder_s_0}
        }
        \subfigure[Finite Size Scaling of the Binder Cumulant $g$][]
        {
            \label{sfig:gettingCrit:collapse_s_0}
            \includegraphics[width=0.47\textwidth]{plots/collapse_s_0}
        }
        \caption[Examples of determining critical temperature and exponents]
        {
            \subref{sfig:gettingCrit:binder_s_0} The Binder cumulant \(g\)
                of an square lattice Ising model (\(\sigma=0\))\\
            \subref{sfig:gettingCrit:collapse_s_0} collapsed by finite
                size scaling
        }
        \label{fig:gettingCrit}
    \end{figure}\\

@@

    expectations are met. Hence 5 values of \(\sigma\) are analyzed.
    The analytically known values for \(\sigma = 0\) and the in \cite{Janke1994}
    examined limit of an random lattice \(\sigma \gtrsim 1\) are natural
    choices.
    The other \(\sigma\) are chosen to represent regions, where the behavior of
    \(T_c\) shows some characteristics. (Namely there is a plateau at small
    \(\sigma\), a steep decline at medium \(\sigma\) and a shallow
    decline at \(\sigma = 0.5\). This will be shown in the next
    chapter in fig. \ref{fig:Tc}\subref{sfig:Tc:RNG}\subref{sfig:Tc:GG}.)

@@

    are matching the expectations. Especially \(\nu\) and \(\gamma\)

@@

    The results are displayed in fig. \ref{fig:susCrossCheck}. The ratios
    determined by this method confirm the values obtained by collapse.

@@

    disordered Ising model is for every \(\sigma\) in the same universality

@@

\subsection{Critical Temperature}
\label{ssec:binderIntersections}
    \(g\) of different system sizes \(L\) which are intersecting at
    \(T_c\) \cite{Binder1981}.
        \caption[Example of a Binder cumulant to determine the critical temperature]
            (\(\sigma=0\)) interpolated with cubic splines
            (the errorbars are too small to see)\\
        \label{fig:gettingCrit:binder_fit_s_0}
    As an example take fig. \ref{fig:gettingCrit:binder_fit_s_0}.
    deviations.\\
    In this section these \(T_c\) on either the RNG or the GG are
    compared. In the following figures, the RNG will always be on the
    left side and the GG on the right side.
    In fig. \ref{fig:Tc} \(T_c\) is plotted over \(\sigma\).
            \label{sfig:Tc:RNG}
            \includegraphics[width=0.45\textwidth]{plots/RNG_Tc}
            \label{sfig:Tc:GG}
            \includegraphics[width=0.45\textwidth]{plots/GG_Tc}
        \caption[Critical Temperature over different disorder parameters]
            Critical temperatures \(T_c\) over different
            disorder parameters \(\sigma\) for
            \subref{sfig:Tc:RNG} the RNG and
            \subref{sfig:Tc:GG} the GG.\\
        \label{fig:Tc}
    One sees that the RNG on the left has generally a lower critical
    temperature than the GG. Also \(T_c\) of the RNG decreases
    monotonically while \(T_c\) of the GG rises at first. Naively one would
    expect, that while changing the displacement of the nodes monotonically,
    the properties of the system will also change monotonically. The
    maximum at \(\sigma \approx 0.15\) on the GG will be discussed later.
    For big \(\sigma\) the displaced nodes approach the limit of randomly
    distributed nodes, hence \(T_c\) is independent of \(\sigma\) for
    \(\sigma \gg 1\).\\
    Another strange property is the jump from \(\sigma = 0\) to \(\sigma > 0\)
    of \(T_c\) on the GG. To understand that, one has to
    consider the influence of the graph properties on the critical
    temperature.\\
    One basic property of a graph is its degree \(K\) -- sometimes
    called \emph{coordination number} -- is defined as the mean count of
    neighbors per node
        K = \frac{\sum_{\avg{i,j}} 1}{\sum_{i} 1}
        \label{eq:degree}
    It is known, that the degree has an impact on the critical Temperature.
    For example the Honeycomb lattice is of degree \(K=3\)
    and there exists an analytic solution \cite{Wannier1945}.
    Its critical temperature is determined by eq. \eqref{eq:exactHCTc}.
        \cosh\brac{\frac{J}{T_c}}=2 \overset{J=1}{\Longrightarrow} T_c \approx 1.52
        \label{eq:exactHCTc}
    This is lower than for the square lattice with degree 4.\\
    If one plots the degree of the graphs at different \(\sigma\) like
    in fig. \ref{fig:Tc_deg}\subref{sfig:deg:RNG}\subref{sfig:deg:GG},
    one recognizes, that \(T_c\) and \(K\) are evidently correlated.
    It seems reasonable to normalize \(T_c\) by the degree of the underlying
    graph. This is done in fig \ref{fig:Tc_deg}\subref{sfig:Tc_norm_deg:RNG}\subref{sfig:Tc_norm_deg:GG}.
    Indeed the normalization eliminates the jump and reduces the
    slope of the rising on the GG. Hence it reduces differences
    between the RNG and the GG.
            \label{sfig:deg:RNG}
            \includegraphics[width=0.45\textwidth]{plots/RNG_deg}
            \label{sfig:deg:GG}
            \includegraphics[width=0.45\textwidth]{plots/GG_deg}
